---
title: "Basic plot example"
author: "Luke Batts"
date: "23 October 2018"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

A short vignette showing how to take the results of a single model and plot survey length frequency distributions along with fitted distributions estimated by the basic model


#Loading data and formatting

WD should be "LFEM/vignettes" folder.
```{r}
getwd()

```
```{r, warning=F , message=F}

#install.packages("gridBase")
library(grid)
library(gridBase)

load("../data/test.Rdata")
load("../data/lfdat_MON.Rdata")

source("../R/plot_lfem.R")

```

Load the length frequency data and the model we've just run.  These can then be input into the `plot.lfem` function.
```{r}

args(plot.lfem)
```
* `model` is the model
* `Lengths` is the same data imput used for the model run(s)
* `Survey.num` allows selection of a particular survey you would like to plot.  Should be the number of the survey if their names were arranged alphabetically.  In this case it is the IE-IGFS but could also be EVHOE(1) or SP-PORC(3) for the anglerfish example.
* `xlimit` allows the xlimit of the length frequency histograms to be set.  Allowing you to focus on the section of the histograms where more data is available.

```{r}

plot.lfem(model=test,Lengths=lfdat,Survey.num=2,xlimit=140)
```

#####If there there is no model and you have all results stored in lists from a sensitivity analysis then just create a list with Mu, Sd and Lambda

```{r,eval=F}

plot.obj<-list(Mu=test$Mu,Sd=test$Sd,Lambda= test$Lambda)

plot.lfem(plot.obj,Lengths=lfdat,Survey.num=2,xlimit=140)
```

#####Note that the cohort progression lines will only be in the right place if plot window dimensions are adjusted before plotting.




 