---
title: "Hierarchical run example"
author: "LB"
date: "08 April 2019"
header-includes:
- \usepackage{pdflscape}
- \newcommand{\blandscape}{\begin{landscape}}
- \newcommand{\elandscape}{\end{landscape}}
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

A short vignette showing how to run the hierarchical LFEM model on an example dataset.  The example dataset in this case is a cleaned up version of white-bellied anglerfish data from the surveys that cover ICES subarea VII and divisions VIII a,b,d stock.

This method allows multiple surveys from different times of year.


#Loading data and formatting

WD should be "LFEM/Vignettes" folder in order to locate .cpp files later
```{r}
getwd()

```
```{r, warning=F , message=F}

#install.packages(c("Matrix","data.table","plyr","reshape","TMB","ggplot2"))
library(TMB) 
library(plyr)
library(reshape)
library(ggplot2)
library(data.table) 
load("../data/lfdat_MON.RData")
```
The dataframe should be formatted as below for the function, with four columns: Survey, Year, Length and RF (raising factor).  The data has been aggregated by year, survey and length so as to compact the dataframe for speed. Observed log-likelihood is the same if individual fish lengths for each haul were used.

```{r,echo=F}
head(lfdat)
```

Objective functions need to be compiled, here is the code to do it. Set root directory for cpp/dll files so the function can find the .dll files.

```{r,warning=F , message=F}

dllroot<-paste0(dirname(getwd()),'/tmb/')
```

```{r,eval=F}
compile(paste0(dllroot,"hier_cL_CSD.cpp"))
compile(paste0(dllroot,"hier_cL_LSD.cpp"))
compile(paste0(dllroot,"hier_cL_CSD_OBSLL.cpp"))
compile(paste0(dllroot,"hier_cL_LSD_OBSLL.cpp"))

compile(paste0(dllroot,"hier_ck_CSD.cpp"))
compile(paste0(dllroot,"hier_ck_LSD.cpp"))
compile(paste0(dllroot,"hier_ck_CSD_OBSLL.cpp"))
compile(paste0(dllroot,"hier_ck_LSD_OBSLL.cpp"))

compile(paste0(dllroot,"hier_yk_CSD.cpp"))
compile(paste0(dllroot,"hier_yk_LSD.cpp"))
compile(paste0(dllroot,"hier_yk_CSD_OBSLL.cpp"))
compile(paste0(dllroot,"hier_yk_LSD_OBSLL.cpp"))
``` 


Source the function
```{r}

source("../R/hier_fun.R")

```
  
#Function arguments and starting parameters

##Survey information
####Values should be entered in alphabetical order of surveys in each case.

How many years of data in each survey?  What year does data for each survey start?   
```{r}
  no.years<-c(14,14,14) 
  year0<-c(2003,2003,2003)
```

age1 is the assumed age of the first component in each of the surveys.  In this case we assume that for the first two surveys alphabetically (EVHOE and IE-IGFS) that the first component observed in the length frequency data is approximately 0.875 years old, using the common assumption fish are born on the 1st of Jan (i.e.  the midpoint of the fourth quarter of the year when these surveys are conducted).  SP-PORC is mainly conducted over september so we set the age1 at 0.73 

```{r}
age1<-c(0.875,0.875,0.73) 
    
```
  
##Starting parameters

* *L* is the mean of the final component

* *l* is the mean of the first component

* *k.reparam* is the starting growth parameter

* sigma.start is the starting standard deviation parameter(s)

* SD.type =
    + 3 -> linear SD relative to means, needs two sigma.start parameters e.g. c(5,10)
    + 4 -> constant SD
    
* RE.type =
    + 1 -> cohort random effects on $l$ and $L$
    + 2 -> cohort random effects on $l$ and $k$
    + 3 -> cohort random effects on $l$ and yearly random effect on $k$
    
* fix.RESD = 
    + TRUE -> standard deviations for the random effects are fixed to input values.  These need to be specified in the function as sdl and sdL or sdk.  This gives models more stability when data is lacking, but still provides enough flexibility (see haddock example of paper).
    + FALSE -> Estimates sd of random effects
 
  \newpage
#Run the function
If rel.tolerance is set at 1e-8 as is standard then this will take some time to converge

```{r,eval=F}
#Example where sd of RE is estimated
hier_test<- hier.LFEM(year0=year0,no.years=no.years,age1=age1,L=130,l=(16),
k.reparam=0.83,sigma.start=c(6,10),No.comp=9,SD.type=3,RE.type=1,fix.RESD=F,Lengths=lfdat,
niter=10000,rel.tolerance=1e-1,dllroot = dllroot)


#Example where sd of RE is fixed by the user
hier_test_fixedRESD<- hier.LFEM(year0=year0,no.years=no.years,age1=age1,L=130,l=(16),
k.reparam=0.83,sigma.start=c(6,10),No.comp=9,SD.type=3,RE.type=1,fix.RESD=T,Lengths=lfdat,
niter=10000,rel.tolerance=1e-1,dllroot = dllroot,sdl=-2,sdL=-2)
```

Load the test object if you haven't run the model
```{r}
load("../data/hier_test.RData")
```
\newpage
#Results
There are many results from the model
```{r}
str(hier_test)
```
First thing to check is the convergence
```{r}
plot(hier_test$obs.llike)

```

Then explore the rest of the results

Mu's and Sd's are now matrices rather than a vector for each survey.  If we take a look at Mu.all.years this displays all component means including those that are used in the unbserved years.  In this case the model assumes nine components(m), so we have (m-1) unobserved years before and after our survey data.  Adding these in allows us to model cohorts.  $l$ values for unbserved years are given the average, as are their corresponding $L$.  This gives the hierrarchical model more stability but also allows us to focus on the variability in the cohorts we observe from their very first year (i.e. the first year of our survey data).

\newpage

```{r,results= 'markup',comment = ""}
hier_test$Mu.all.years[,,1]
```

We can plot the results with same function as we did the basic example.   
```{r}
source("../R/plot_lfem.R")
library(grid)
library(gridBase)
plot.lfem(model=hier_test,Lengths=lfdat,Survey.num=2,xlimit=140)
```




